---
title: "Linear_regression"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linear_regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
   collapse = TRUE,
  comment = "#>"
)
```

<style>
body {
text-align: justify}
</style>
The objective of this package is to compute several characteristics of a
**Linear Regression**, like: Coefficients, Residual, fitted values, degrees of
freedom, t-values, etc. Given, by the user, the arguments formula and data.
*Formula* is an argument of class formula, which contains the
Y(*dependent*) and X´s(*independent*) variables.
*Data*is an argument of class data.frame, which contains the data required to
make all the matrix calculations and return the characteristics.


The information below is presented trough the following S3 methods/functions:


**print()**

Print method prints the coefficients and coefficients names:

```{r setup} 
library(Lab04)
data("iris")
model_object<-linreg(Petal.Length~Species, data=iris)
print(model_object)
```

**plot()**

It returns two plots: 

1. The mean and the median results of the **Residuals vs Fitted Values**.

1. The mean and the median results of the Standardized Residuals vs Fitted Values. **Scale-Location**

```{r}
plot(model_object)
```

**summary()**

Ir returns a summary of the main characteristics of the *Linear Regression*, presented as a Data.frame class and one strings, using the function cat().

```{r}
summary(model_object)
```

**resid()**
It returns the vector of residuals e_hat
```{r}
resid(model_object)
```

**pred()**
It returns the predicted values y_hat
```{r}
pred(model_object)
```

**coef()**
It returns the coefficients of the Beta matrix.
```{r}
coef(model_object)
```
