---
title: "Linear_regression"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linear_regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
   collapse = TRUE,
  comment = "#>"
)
```

<style>
body {
text-align: justify}
</style>
The objective of this package is to compute several characteristics of a
**Linear Regression**, like: Coefficients, Residual, fitted values, degrees of
freedom, t-values, etc. It computes and save all the results trough the function call
**linreg**.

**linreg** has two arguments *linreg(formula=Petal.Length~Species, data=iris)*

*Formula* is an argument of class formula, which contains the
Y(*dependent*) and X´s(*independent*) variables.
*Data* is an argument of class data.frame, which contains the data required to
make all the matrix calculations and return the characteristics.

```{r setup}
library(Lab04)
```


The information below is presented trough the following S3 methods/functions:


**print()**

Print method prints the model call and estimated coefficients:

```{r} 
data("iris")
linreg(Petal.Length~Species, data=iris)
model_object<-linreg(Petal.Length~Species, data=iris)
print(model_object)
```

**plot()**

It returns two plots for model diagnostics: 

1. **Residuals vs Fitted Values**. Estimated mean and median residuals for predicted values are indicated in the plot and joined by lines. 

1. **Scale-Location**. Squared absolute standardized residuals vs Fitted Values. 
Estimated mean and median Squared absolute standardized residuals for predicted 
values are indicated in the plot and joined by lines.

```{r fig1, fig.height = 5, fig.width = 7, fig.align = "center"}
plot(model_object)
```

**summary()**

Ir returns a summary of the main characteristics of the *Linear Regression*, presented as a Data.frame class and one strings, using the function cat().

```{r}
summary(model_object)
```

**resid()**
It returns the vector of residuals e_hat
```{r}
resid(model_object)
```

**pred()**
It returns the predicted values y_hat
```{r}
pred(model_object)
```

**coef()**
It returns the regression coefficients.
```{r}
coef(model_object)
```
