---
title: "Nycflights13"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nycflights13}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Lab04)
library(dplyr)
library(caret)
library(stats)
library(nycflights13)
library(ggplot2)
library(GGally)
```
In this vignette, data from the nycflights13 package is modelled. The data concerns 
flights that departed NYC in 2013. Here we are using ridge regression to predict arrival
delay time, using data from the flights dataset as well as the weather dataset. The caret
package is used to divide the data into training, validation and test set, as well as
imputing missing values. The predictions are made using the ridgereg function in the 
package Lab04.
## Load data
```{r}
data(flights)
data(weather)
```
## Number of variables and observations
```{r}
dim(flights)
dim(weather)
```
## View data
```{r}
head(flights)
head(weather)

names(flights)
names(weather)
```
## Inspect missing
```{r}
missing_flights <- flights %>% summarise_all(funs(sum(is.na(.))))

(missing_flights_perc <- round(100*(missing_flights/nrow(flights)),2))

missing_weather <- weather %>% summarise_all(funs(sum(is.na(.))))

(missing_weather_perc <- round(100*(missing_weather/nrow(weather)),2))

```
Summary:
- For flights data, max 2.8% of observations missing\n
- For weather data, 80% missing for wind_gust, and 10.5% for pressure\n
## Exclude observations with missing arr_delay
```{r}
flights_1a <- flights%>% filter(!is.na(arr_delay))
```
## Select variables
```{r}
#- Create weekday variable
flights_2a <- mutate(flights_1a,weekday=weekdays(as.Date(substring(time_hour,1,10))))

#- Remove variables:

flights_2b <- select(flights_2a,-year,-day,-dep_time,-arr_time,-minute,-sched_dep_time)
weather_1a <- select(weather,-year,-month,-day,-hour,-wind_gust)
```
## Join datasets
```{r}
flights_3a <- left_join(flights_2b, weather_1a,
                        by = c("origin" = "origin", "time_hour"="time_hour"))

#- Remove time_hour variable
flights_3b <- select(flights_3a,-time_hour)

```
## Handle type of variable factor
```{r}
flights_4a <- mutate(flights_3b,
                     flight = factor(flight),
                     month = factor(month),
                     weekday = factor(weekday),
                     hour = factor(hour),
                     carrier = factor(carrier),
                     tailnum = factor(tailnum),
                     origin = factor(origin),
                     dest = factor(dest))
```
## Variable distributions, values
```{r}
summary(flights_4a)

#- Number of factor levels
str(select(flights_4a,month,  weekday, hour, carrier, flight, tailnum, origin,dest))

#- Scatter plot matrices

set.seed(42)
data_scatter <- sample_n(flights_4a,1000)

ggpairs((select(data_scatter, arr_delay,dep_delay,air_time,distance)))

ggpairs((select(data_scatter, temp,dewp,humid,wind_speed,precip,pressure,visib)))

#- Boxplots
ggplot(data_scatter, aes(x = hour, y = arr_delay)) +
  geom_boxplot(alpha = 0.1, aes(group = hour)) +
  xlab("Scheduled hour of departure") + ylab("Arrival delay")

ggplot(data_scatter, aes(x = weekday, y = arr_delay)) +
  geom_boxplot(alpha = 0.1, aes(group = weekday)) +
  xlab("Weekday of departure") + ylab("Arrival delay")

ggplot(data_scatter, aes(x = month, y = arr_delay)) +
  geom_boxplot(alpha = 0.1, aes(group = month)) +
  xlab("Month") + ylab("Arrival delay")

ggplot(data_scatter, aes(x = origin, y = arr_delay)) +
  geom_boxplot(alpha = 0.1, aes(group = origin)) +
  xlab("Origin") + ylab("Arrival delay")

ggplot(data_scatter, aes(x = dest, y = arr_delay)) +
  geom_boxplot(alpha = 0.1, aes(group = dest)) +
  xlab("Destination") + ylab("Arrival delay")

ggplot(data_scatter, aes(x = carrier, y = arr_delay)) +
  geom_boxplot(alpha = 0.1, aes(group = carrier)) +
  xlab("Carrier") + ylab("Arrival delay")

#- Tabulate
ggplot(data_scatter, aes(x=temp)) + geom_histogram()
ggplot(data_scatter, aes(x=dewp)) + geom_histogram()
ggplot(data_scatter, aes(x=humid)) + geom_histogram()
ggplot(data_scatter, aes(x=wind_speed)) + geom_histogram()
ggplot(data_scatter, aes(x=precip)) + geom_histogram()
ggplot(data_scatter, aes(x=pressure)) + geom_histogram()
ggplot(data_scatter, aes(x=visib)) + geom_histogram()
```
## Investigate missing values and filter if too many missing
```{r}
flights_5a <- flights_4a
flights_5a$perc_missing_row <-
  apply(flights_5a, 1, function(x) round(100*(sum(is.na(x))/length(x)),2))

flights_5a$perc_missing_row_e_p <-
  apply(select(flights_5a,-pressure), 1,
        function(x) round(100*(sum(is.na(x))/length(x)),2))


table(flights_5a$perc_missing_row)

table(flights_5a$perc_missing_row_e_p)

flights_5b <- filter(flights_5a,perc_missing_row<5)

table(flights_5b$perc_missing_row)

table(flights_5b$perc_missing_row_e_p)

#- Columns
missing_cols <- flights_5b %>% summarise_all(funs(sum(is.na(.))))

(missing_cols_perc <- round(100*(missing_cols/nrow(flights)),2))
```
## Investigate dest and carrier variables
```{r}
select_dest <- filter(flights_5b,dest=="LEX")

by_dest <- dplyr::group_by(flights_5b, dest)

count_dest <- dplyr::summarise(by_dest,
                               count_dest = n())

by_carrier <- dplyr::group_by(flights_5b, carrier)

count_carrier <- dplyr::summarise(by_carrier,
                                  count_carrier = n(),
                                  count_missing = sum(is.na(arr_delay)),
                                  perc_missing = (count_missing/count_carrier)*100,
                                  delay_mean = mean(arr_delay, na.rm = TRUE),
                                  .groups = "keep")

arrange(count_dest,count_dest)
arrange(count_carrier,count_carrier)


```
## Filter on dest > 100 counts
```{r}
flights_6a <- left_join(flights_5b, count_dest, by = c("dest" = "dest"))

flights_6b <- filter(flights_6a,count_dest>=100)

flights_6c <- mutate(flights_6b,dest = factor(dest))

```

## Divide into training, validation and test
```{r}
#- Set data for modelling

data <- flights_6c

set.seed(42)
data <- sample_n(data,80000)

data <- select(data, arr_delay, air_time, month, origin, dep_delay,
               distance, hour, humid, temp,  wind_speed,
               precip,visib,dewp,carrier,weekday, pressure, dest) #dest

set.seed(42)
inTrainVal <- caret::createDataPartition(y = data$arr_delay, p = .95,list = FALSE)
train_val <- data[ inTrainVal,]
testing <- data[-inTrainVal,]

nrow(train_val)/nrow(data)
nrow(testing)/nrow(data)

set.seed(42)
train_p <- 0.8/0.95
inTrain <- caret::createDataPartition(y = train_val$arr_delay, p = train_p,list = FALSE)

training <- train_val[ inTrain,]
validation <- train_val[-inTrain,]

nrow(training)/nrow(data)
nrow(validation)/nrow(data)
nrow(testing)/nrow(data)

#- Convert to data frames
training <- as.data.frame(training)
validation <- as.data.frame(validation)
testing <- as.data.frame(testing)

```
## Impute missing values
```{r}

set.seed(42)

missing_model = preProcess(select(training,-arr_delay), "knnImpute")
training_x_imp <- predict(missing_model, select(training,-arr_delay))
validation_x_imp <- predict(missing_model, select(validation,-arr_delay))
testing_x_imp <- predict(missing_model, select(testing,-arr_delay))

training <- cbind(select(training,arr_delay),training_x_imp)
validation <- cbind(select(validation,arr_delay),validation_x_imp)
testing <- cbind(select(testing,arr_delay),testing_x_imp)

```


## Specify model
```{r}
formula <- arr_delay ~ . + precip*wind_speed
training <- as.data.frame(training)
grid_vec <- c(0,0.2,0.5,0.75,1,1.5,2,2.5,5)

```
## Fit models on training set for different lambda
```{r}
validation_error <- vector(mode="numeric",length=length(grid_vec))

validation_x <- data.frame(model.matrix(object=formula, data=validation)[,-1])

testing_x <- data.frame(model.matrix(object=formula, data=testing)[,-1])
validation_y <- as.vector(unlist(validation[,all.vars(expr=formula)[1]]))
testing_y <- as.vector(unlist(testing[,all.vars(expr=formula)[1]]))

#- Change . to : to match with output from the ridge model
names(validation_x) <- gsub("\\.",":", names(validation_x))
names(testing_x) <- gsub("\\.",":", names(testing_x))

for (i in seq_along(grid_vec)) {
  lambda <- grid_vec[i]
  rr_res_training <- ridgereg(formula=formula, data = training, lambda=lambda)
  validation_error[i] <- rmse(validation_y-predict(rr_res_training,validation_x))
}

title <- c("RMSE validation set knn imputation")
ylab <- c("RMSE")
xlab <- c("lambda")
df_selection <- data.frame(lambda=grid_vec,validation_error=validation_error)
p1 <- ggplot(df_selection, aes(lambda, validation_error))+geom_point()+
  labs(title=title, y=ylab, x=xlab)
p1
```
## Plot RMSE for the different lambda values
```{r}
title <- c("RMSE validation set knn imputation")
ylab <- c("RMSE")
xlab <- c("lambda")
df_selection <- data.frame(lambda=grid_vec,validation_error=validation_error)
p1 <- ggplot(df_selection, aes(lambda, validation_error))+geom_point()+
  labs(title=title, y=ylab, x=xlab)
p1
```
\n
## Calculate RMSE for selected lambda on test set
Use training and validation set to construct the final model
```{r}
lambda <- 1.5
train_val <- rbind(training,validation)
rr_res_final <- ridgereg(formula, data = train_val,lambda=lambda)

test_error <- rmse(testing_y-predict(rr_res_final,testing_x))
print(test_error)

```

